---
title: "Evaulation of DOGE Data"
author: "Peter Ashley"
date: "5/8/25"
output: github_document
---

```{r libraries}
library(tidyverse)
library(rvest)
library(dplyr)
library(httr)
library(jsonlite)
```

# Question: How did DOGE create their "value" property in their API?

# Setting up our data

## Read FPDS Data
Takes over from where I left off at the end of ProjExp.Rmd

```{r read-fpds-data}
fpds <- read_csv("fpds.csv")
fpds <- fpds[-c(1)]
```

## Get AWARD-specific data, ignoring IDV values

```{r filter-award-type}
awards <- fpds %>%
  filter(contract_type == "AWARD")
```

## Convert numeric data to numeric type

```{r convert-numeric}
awards <- awards %>%
  mutate( #Removes dollar signs and commas from money rows, with as.numeric wrap
    obligated_amount = as.numeric(gsub('\\$|,', '', obligated_amount)),
    total_obligated_amount = as.numeric(gsub('\\$|,', '', total_obligated_amount)),
    base_and_exercised_options_value = as.numeric(gsub('\\$|,', '', base_and_exercised_options_value)),
    total_base_and_exercised_options_value = as.numeric(gsub('\\$|,', '', total_base_and_exercised_options_value)),
    ultimate_contract_value = as.numeric(gsub('\\$|,', '', ultimate_contract_value)),
    total_ultimate_contract_value = as.numeric(gsub('\\$|,', '', total_ultimate_contract_value)),
    fees_paid_for_use_of_service = as.numeric(gsub('\\$|,', '', fees_paid_for_use_of_service))
  ) %>%
  mutate_if(is.numeric, ~replace(., is.na(.), 0))
```

Sourced from https://stackoverflow.com/questions/8161836/how-do-i-replace-na-values-with-zeros-in-an-r-dataframe#:~:text=mutate_if(is.numeric%2C%20~replace(.%2C%20is.na(.)%2C%200))

## Select necessary columns, rename them for shortening

```{r awards-numeric}
awards_num <- awards %>%
  select(piid, idv_piid, value, savings, obligated_amount, total_obligated_amount, base_and_exercised_options_value, total_base_and_exercised_options_value, ultimate_contract_value, total_ultimate_contract_value, fees_paid_for_use_of_service) %>%
  rename(obligated = obligated_amount, total_obligated = total_obligated_amount, options = base_and_exercised_options_value, total_options = total_base_and_exercised_options_value, ucv = ultimate_contract_value, total_ucv = total_ultimate_contract_value, fees = fees_paid_for_use_of_service)
```

## Create Calculated Columns that are repeatedly used for future calculations

```{r basic-calculated-columns}
awards_num <- awards_num %>%
  mutate(
    obl_totalobl = (obligated + total_obligated),
    obl_totalobl_abs = (abs(obligated) + abs(total_obligated)),
    opt_totalopt = (options + total_options),
    opt_totalopt_abs = (abs(options) + abs(total_options)),
    ucv_totalucv = (ucv + total_ucv),
    ucv_totalucv_abs = (abs(ucv) + abs(total_ucv))
  )
awards_num
```

# Initial Check-Values

### Important Variables and Concepts:
#### checks: the dataframe, named because it holds a large number of check-values
#### value: what we are solving for, from the doge api
#### minval: the lowest approximation of a row's value based on other variables
#### obligated: one of the major contributors to value, fpds
#### options: one of the major contributors to value, fpds
#### ultimate_contract_value (ucv): one of the major contributors to value, fpds
#### fees: another number that contributes to value (included later), fpds
#### savings: another number from doge that contributes to value
#### total: prefix for obligated, options, and ucv columns
#### previous: prefix for obligated, options, and ucv columns (unused, no values)

```{r check-values}
checks <- awards_num %>%
  mutate(
    a = abs(value - total_obligated),
    a1 = abs(value - total_obligated - savings),
    a2 = abs(value - total_obligated + savings),
    b = abs(value - obl_totalobl),
    b1 = abs(value - obl_totalobl - savings),
    b2 = abs(value - obl_totalobl + savings),
    c = abs(value - obl_totalobl_abs),
    c1 = abs(value - obl_totalobl_abs - savings),
    c2 = abs(value - obl_totalobl_abs + savings),
    d = abs(value - total_options),
    d1 = abs(value - total_options - savings),
    d2 = abs(value - total_options + savings),
    e = abs(value - opt_totalopt),
    e1 = abs(value - opt_totalopt - savings),
    e2 = abs(value - opt_totalopt + savings),
    f = abs(value - opt_totalopt_abs),
    f1 = abs(value - opt_totalopt_abs - savings),
    f2 = abs(value - opt_totalopt_abs + savings),
    g = abs(value - total_ucv),
    g1 = abs(value - total_ucv - savings),
    g2 = abs(value - total_ucv + savings),
    h = abs(value - ucv_totalucv),
    h1 = abs(value - ucv_totalucv - savings),
    h2 = abs(value - ucv_totalucv + savings),
    i = abs(value - ucv_totalucv_abs),
    i1 = abs(value - ucv_totalucv_abs - savings),
    i2 = abs(value - ucv_totalucv_abs + savings)
  )
checks
```
## Takes all rows and gets the Minimum Value from all columns

```{r get-all-minval}
checks <- checks %>%
  rowwise() %>% # dplyr rowwise() used here to take min() on row instead of col
  mutate(
    minval = min(c_across(a:i2))
  ) %>%
  ungroup() # ungroup the rowwise() to prevent any accidental errors
checks
```

## Checks which column the Minimum Value is first seen in

```{r get-mincol}
checks <- checks %>%
  mutate(
    mincol = # Huge if statement but it works
      if_else(a == minval, "a",
      if_else(a1 == minval, "a1",
      if_else(a2 == minval, "a2",
      if_else(b == minval, "b",
      if_else(b1 == minval, "b1",
      if_else(b2 == minval, "b2",
      if_else(c == minval, "c",
      if_else(c1 == minval, "c1",
      if_else(c2 == minval, "c2",
      if_else(d == minval, "d",
      if_else(d1 == minval, "d1",
      if_else(d2 == minval, "d2",
      if_else(e == minval, "e",
      if_else(e1 == minval, "e1",
      if_else(e2 == minval, "e2",
      if_else(f == minval, "f",
      if_else(f1 == minval, "f1",
      if_else(f2 == minval, "f2",
      if_else(g == minval, "g",
      if_else(g1 == minval, "g1",
      if_else(g2 == minval, "g2",
      if_else(h == minval, "h",
      if_else(h1 == minval, "h1",
      if_else(h2 == minval, "h2",
      if_else(i == minval, "i",
      if_else(i1 == minval, "i1",
      if_else(i2 == minval, "i2",
      "ERROR"
      ))))))))))))))))))))))))))
    )
  )
```

## Looks for rows which were affected by PlusSavings functions

```{r evaluate-plus-savings}
checks %>%
  filter(savings > 0) %>%
  select(piid,a2,b2,c2,d2,e2,f2,g2,h2,i2,mincol) %>%
  rowwise() %>%
  mutate(
    minval = min(a2,b2,c2,d2,e2,f2,g2,h2,i2)
  ) %>%
  ungroup() %>%
  filter(minval < 1)
```

## Looks exclusively for rows with a PlusSavings column for a mincol
This indicates no other columns likely impacted it, 
can go back and check for each individual row

```{r evaluate-plus-savings-deep}
checks %>%
  filter(savings > 0) %>%
  select(piid,a2,b2,c2,d2,e2,f2,g2,h2,i2,mincol) %>%
  mutate(
    minval = min(a2,b2,c2,d2,e2,f2,g2,h2,i2)
  ) %>%
  filter(minval < 1) %>%
  filter(mincol %in% c("a2","b2","c2","d2","e2","f2","g2","h2","i2"))
```

## `r checks %>% filter(abs(minval) > 1) %>% count()` values have no solved formula (minval < 1), can we do better?

```{r rows-no-formula}
checks %>%
  filter(abs(minval) > 1) %>%
  count()
```

# An In-Depth expansive check

## Generate a string of all letter/number combos a-l, 1-9

```{r get-minval-string}
s <- ""
for (letter in c("a","b","c","d","e","f","g","h","i","k","l")){
  for (i in 1:9){
    s <- paste0(s, letter, i, ",")
  }
}
s
```

## Expands from a-i to a-l, and from 3 to 9 formulas, incorporating fees

```{r create-check-columns}
checks <- awards_num %>%
  mutate(
    a1 = abs(value - total_obligated),
    a2 = abs(value - total_obligated - fees),
    a3 = abs(value - total_obligated + fees),
    a4 = abs(value - total_obligated - savings),
    a5 = abs(value - total_obligated + savings),
    a6 = abs(value - total_obligated - savings + fees),
    a7 = abs(value - total_obligated - savings - fees),
    a8 = abs(value - total_obligated + savings + fees),
    a9 = abs(value - total_obligated + savings - fees),
    b1 = abs(value - obl_totalobl),
    b2 = abs(value - obl_totalobl - fees),
    b3 = abs(value - obl_totalobl + fees),
    b4 = abs(value - obl_totalobl - savings),
    b5 = abs(value - obl_totalobl + savings),
    b6 = abs(value - obl_totalobl - savings + fees),
    b7 = abs(value - obl_totalobl - savings - fees),
    b8 = abs(value - obl_totalobl + savings + fees),
    b9 = abs(value - obl_totalobl + savings - fees),
    c1 = abs(value - obl_totalobl_abs),
    c2 = abs(value - obl_totalobl_abs - fees),
    c3 = abs(value - obl_totalobl_abs + fees),
    c4 = abs(value - obl_totalobl_abs - savings),
    c5 = abs(value - obl_totalobl_abs + savings),
    c6 = abs(value - obl_totalobl_abs - savings + fees),
    c7 = abs(value - obl_totalobl_abs - savings - fees),
    c8 = abs(value - obl_totalobl_abs + savings + fees),
    c9 = abs(value - obl_totalobl_abs + savings - fees),
    d1 = abs(value - total_options),
    d2 = abs(value - total_options - fees),
    d3 = abs(value - total_options + fees),
    d4 = abs(value - total_options - savings),
    d5 = abs(value - total_options + savings),
    d6 = abs(value - total_options - savings + fees),
    d7 = abs(value - total_options - savings - fees),
    d8 = abs(value - total_options + savings + fees),
    d9 = abs(value - total_options + savings - fees),
    e1 = abs(value - opt_totalopt),
    e2 = abs(value - opt_totalopt - fees),
    e3 = abs(value - opt_totalopt + fees),
    e4 = abs(value - opt_totalopt - savings),
    e5 = abs(value - opt_totalopt + savings),
    e6 = abs(value - opt_totalopt - savings + fees),
    e7 = abs(value - opt_totalopt - savings - fees),
    e8 = abs(value - opt_totalopt + savings + fees),
    e9 = abs(value - opt_totalopt + savings - fees),
    f1 = abs(value - opt_totalopt_abs),
    f2 = abs(value - opt_totalopt_abs - fees),
    f3 = abs(value - opt_totalopt_abs + fees),
    f4 = abs(value - opt_totalopt_abs - savings),
    f5 = abs(value - opt_totalopt_abs + savings),
    f6 = abs(value - opt_totalopt_abs - savings + fees),
    f7 = abs(value - opt_totalopt_abs - savings - fees),
    f8 = abs(value - opt_totalopt_abs + savings + fees),
    f9 = abs(value - opt_totalopt_abs + savings - fees),
    g1 = abs(value - total_ucv),
    g2 = abs(value - total_ucv - fees),
    g3 = abs(value - total_ucv + fees),
    g4 = abs(value - total_ucv - savings),
    g5 = abs(value - total_ucv + savings),
    g6 = abs(value - total_ucv - savings + fees),
    g7 = abs(value - total_ucv - savings - fees),
    g8 = abs(value - total_ucv + savings + fees),
    g9 = abs(value - total_ucv + savings - fees),
    h1 = abs(value - ucv_totalucv),
    h2 = abs(value - ucv_totalucv - fees),
    h3 = abs(value - ucv_totalucv + fees),
    h4 = abs(value - ucv_totalucv - savings),
    h5 = abs(value - ucv_totalucv + savings),
    h6 = abs(value - ucv_totalucv - savings + fees),
    h7 = abs(value - ucv_totalucv - savings - fees),
    h8 = abs(value - ucv_totalucv + savings + fees),
    h9 = abs(value - ucv_totalucv + savings - fees),
    i1 = abs(value - ucv_totalucv_abs),
    i2 = abs(value - ucv_totalucv_abs - fees),
    i3 = abs(value - ucv_totalucv_abs + fees),
    i4 = abs(value - ucv_totalucv_abs - savings),
    i5 = abs(value - ucv_totalucv_abs + savings),
    i6 = abs(value - ucv_totalucv_abs - savings + fees),
    i7 = abs(value - ucv_totalucv_abs - savings - fees),
    i8 = abs(value - ucv_totalucv_abs + savings + fees),
    i9 = abs(value - ucv_totalucv_abs + savings - fees),
    j1 = abs(value - abs(obligated)),
    j2 = abs(value - abs(obligated) - fees),
    j3 = abs(value - abs(obligated) + fees),
    j4 = abs(value - abs(obligated) - savings),
    j5 = abs(value - abs(obligated) + savings),
    j6 = abs(value - abs(obligated) - savings + fees),
    j7 = abs(value - abs(obligated) - savings - fees),
    j8 = abs(value - abs(obligated) + savings + fees),
    j9 = abs(value - abs(obligated) + savings - fees),
    k1 = abs(value - abs(options)),
    k2 = abs(value - abs(options) - fees),
    k3 = abs(value - abs(options) + fees),
    k4 = abs(value - abs(options) - savings),
    k5 = abs(value - abs(options) + savings),
    k6 = abs(value - abs(options) - savings + fees),
    k7 = abs(value - abs(options) - savings - fees),
    k8 = abs(value - abs(options) + savings + fees),
    k9 = abs(value - abs(options) + savings - fees),
    l1 = abs(value - abs(ucv)),
    l2 = abs(value - abs(ucv) - fees),
    l3 = abs(value - abs(ucv) + fees),
    l4 = abs(value - abs(ucv) - savings),
    l5 = abs(value - abs(ucv) + savings),
    l6 = abs(value - abs(ucv) - savings + fees),
    l7 = abs(value - abs(ucv) - savings - fees),
    l8 = abs(value - abs(ucv) + savings + fees),
    l9 = abs(value - abs(ucv) + savings - fees)
  )
```

## Find minval

```{r minval-large}
checks <- checks %>%
  rowwise() %>%
  mutate(
    minval = min(c_across(a1:l9))
  ) %>%
  ungroup()
checks
```



```{r rows-no-formula-2}
checks %>%
  filter(abs(minval) > 1) %>%
  count()

checks %>%
  filter(abs(minval) > 1) %>%
  ggplot(mapping = aes(minval)) +
  geom_histogram(binwidth = 1000000)
```

## Lots of high-skewed outliers, removed everything over 2.5m
This chart shows awards that I couldn't find a formula for,
with most values closer to 0, but some ranging incredibly high

```{r high-value-histogram}
checks %>%
  filter(abs(minval) > 1) %>%
  filter(abs(minval) < 2500000) %>%
  count()

checks %>%
  filter(abs(minval) > 1) %>%
  filter(abs(minval) < 2500000) %>%
  ggplot(mapping = aes(minval)) +
  geom_histogram(binwidth=100000)
```

## Zooming in further, sub-100k
This chart is further zoomed in, but doesn't really show much.

```{r low-value-histogram}
checks %>%
  filter(abs(minval) > 1) %>%
  filter(abs(minval) < 100000) %>%
  count()

checks %>%
  filter(abs(minval) > 1) %>%
  filter(abs(minval) < 100000) %>%
  ggplot(mapping = aes(minval)) +
  geom_histogram(binwidth=2500)
```

# Evaluating lowest check-values

## Convert checks from wide-form to long-form data

```{r checks-long}
checks_long <- checks %>%
  pivot_longer(cols="a1":"l9",
               names_to="check_column",
               values_to="check_value")
checks_long %>%
  filter(check_value < 1)
```

# Most effective formula-variable combinations

```{r effective-formula-variable-combos}
checks_long %>%
  filter(check_value < 1) %>%
  group_by(check_column) %>%
  count() %>%
  rename(formula = check_column) %>%
  arrange(desc(n)) %>%
  head(9)
```

###Most effective combos:
####i1 = abs(value - ucv_totalucv_abs), +- fees
####g1 = abs(value - total_ucv), +- fees
####h1 = abs(value - ucv_totalucv), +- fees

This shows that the ucv variable is one of the more critical ones

## Separate variable and formula from check-column

```{r separate-check-column}
checks_long <- checks_long %>%
  mutate(
    var = substr(check_column, 1, 1),
    formula = as.numeric(substr(check_column, 2, 2))
  )
checks_long
```

## Evaluating variables

```{r effective-variables}
var_eval <- checks_long %>%
  filter(check_value < 1) %>%
  group_by(var) %>%
  count() %>%
  arrange(desc(n))
var_eval
var_eval %>%
  ggplot(mapping=aes(x=var, y=n)) + 
    geom_point()
```

In evaluating variables, we find the following:

Most effective:
i: ucv_totalucv_abs
g: total_ucv
c: obl_totalobl_abs

Least effective:
l: ucv
k: options
j: obligated

From this, we see that calculated columns with absolute values tend to do well, 
and the columns which have few non-blank values do poorly. This is expectable.

## Evaluating formulae

```{r effective-formulae}
formulae_eval <- checks_long %>%
  filter(check_value < 1) %>%
  group_by(formula) %>%
  count() %>%
  arrange(desc(n))
formulae_eval
formulae_eval %>%
  ggplot(mapping=aes(x=formula, y=n)) + 
    geom_point()
```

In evaluating effectiveness of formulae, we find the following:

Most effective:
4: abs(value - var - savings)
6: abs(value - var - savings + fees)
7: abs(value - var - savings - fees)

Least effective:
5 = abs(value - var + savings)
8 = abs(value - var + savings + fees)
9 = abs(value - var + savings - fees)

This was more interesting, as it showed which formulae actually worked.
The fact that 13000 still pass surprises me though

## Evaluating formulae with savings

```{r effective-formulae-savings}
form_savings <- checks_long %>%
  filter(check_value < 1) %>%
  filter(savings > 0) %>%
  group_by(formula) %>%
  count() %>%
  arrange(desc(n))
form_savings
form_savings %>%
  ggplot(mapping=aes(x=formula, y=n)) + 
    geom_point()
```

When only rows with a non-zero savings value are used, 
formulas 5, 8, and 9 become obsolete

